`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2020-02-05 analysis/paper/sim_survey_paper.Rmd
Revise first two paragraphs of intro -> new draft of intro

Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies. While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998] and the quality of this information depends on surveys and analyses that maximize information while minimizing the expense of data collection. While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent. Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect sufficient biological sub-samples (e.g. lengths and ages) to represent the entire target population [@Pennington1994; @Pennington2002; @Stewart2014]. This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015]. These complexities may explain why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively. The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure). In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.

@@

While there are many model-based options for obtaining an abundance index from survey data [e.g. @Thorson2015], design-based approaches, such as stratified analyses, are often used. Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$). Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:

@@

In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)). Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider `range`) and exhibits stronger correlation across ages than years (i.e. lower `phi_year` and higher `phi_age`). Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure. Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly such that values from one function call are forwarded to the next function (i.e. output from the two calls below are functionally the same though the approach is slightly different).

@@

The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs [e.g. @Cochran1977]. Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates [e.g. @smith1990; @Berg2014; @Thorson2015]. **`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.). Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested. Existing and future approaches for calculating age-based indices of abundance can be simulation tested using **`SimSurvey`**.

@@

Assessing ages for a large number of fish is very time-consuming and, as such, length-stratified sampling is often used to estimate age frequencies of fish populations. The resultant sub-sample is used to construct an age-length key (i.e. the probability a fish is a specific age given length) and age frequencies are obtained by applying this key to length-frequencies obtained via more expansive random sampling. One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012]. Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g. @Berg2012].

@@

The negative consequences of excessive length sampling on abundance at age estimates are most apparent under the low set density scenarios. The precision of the estimates are also the poorest when fewer sets are conducted. Sets, however, are the most costly sample to obtain and, as such, the set densities of fisheries-independent surveys tend to be on the lower end of the scenarios presented in this paper. The set density of the multi-species survey conducted by Fisheries and Oceans Canada in the Newfoundland region, for instance, ranges between 0.001 and 0.002 sets / km^2^. Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively easy to obtain. Again, using Fisheries and Oceans Canada sampling protocol from the Newfoundland region as an example, it is not uncommon for length sampling effort to be capped at 500 measurements per set. The results presented in this paper paradoxically suggest that better abundance at age estimates are obtained by lowering the length sampling cap. Results also indicate that reductions to length sampling effort would not significantly impact abundance at length estimates. Nonetheless, it is hard to be prescriptive because this simulation is far from a perfect reflection of reality and it focuses on one case study. Further research is required on the consequences of some of the assumptions of this simulation as well as the interaction between age-specific clustering, length-stratified sampling and the age-length key. Ideally, the simulation would also include multiple species with different life-histories and distributions, and also integrate a cost component to assess the trade-offs between information and cost. Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key [e.g. @Berg2012].

@@

Standard notation for use in the analysis of stratified-random survey data [modified from @Smith1981]:

>>>>

Good survey design is a critical prerequisite for obtaining accurate and precise estimates of a population of interest. The need for sound sampling techniques is as relevant today as it was in the 1930's when much of the theory and principles behind the collection of scientific data were laid by statisticians such as R.A. Fisher and Jerzy Neyman [@fienberg1996]. Then and now, we are experiencing a period of increasing human activity and robust methods are required to detect the impacts of our actions on the natural world [@nichols2006]. Survey quality is especially important for conservation-oriented science as monitoring data are used to identify species at risk of extinction [@martin2007], to assess the efficacy of recovery plans [@campbell2002], and to determine precautionary levels of exploitation [@sutherland2001]. Surveys with a solid basis in sampling theory have therefore become a core component of many ecological monitoring programs throughout the world.

One of the biggest challenges associated with surveys is the costs of planning and implementing effective and efficient sampling designs [@reynolds2011]. Despite the costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998]. For instance, large population declines can be overlooked by relying solely on catch and effort data from commercial fisheries [@rose1999]. Such difficulties have contributed to the growing reliance on fisheries-independent surveys to monitor fish populations. Planning or improving such surveys is an important but non-trivial task because of the inherent complexity of both the survey and target population. Specifically, fisheries-independent surveys usually involve multiple levels of sampling and fish populations often show size-structured spatial and temporal dynamics [@Nelson2014]. These features are not easily captured by simple statistical models, which makes it difficult to determine optimal sampling and sub-sampling schemes [@Pennington1994; @Pennington2002; @Stewart2014]. In such situations it is common for practitioners to resort to simulations as a means of testing the efficacy of various sampling designs (e.g. [@thomas2010]), however, the complexities associated with fisheries-independent surveys may explain why simulations that test the full sampling and analytical pathway are rare (but see [@puerta2019; @schnute2003]).

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) (here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015]) and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively. The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure). In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.

@@

While there are many model-based options for obtaining an abundance index from survey data (e.g. [@Thorson2015]), design-based approaches, such as stratified analyses, are often used. Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$). Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:

@@

In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)). Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider `range`) and exhibits stronger correlation across ages than years (i.e. lower `phi_year` and higher `phi_age`). Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure. Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe (`%>%`; [@bache2014]) friendly such that values from one function call are forwarded to the next function (i.e. output from the two calls below are functionally the same though the approach is slightly different).

@@

The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs (e.g. [@Cochran1977]). Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates (e.g. [@smith1990; @Berg2014; @Thorson2015]). **`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.). Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested. Existing and future approaches for calculating age-based indices of abundance can be simulation tested using **`SimSurvey`**.

@@

Assessing ages for a large number of fish is very time-consuming and, as such, length-stratified sampling is often used to estimate age frequencies of fish populations. The resultant sub-sample is used to construct an age-length key (i.e. the probability a fish is a specific age given length) and age frequencies are obtained by applying this key to length-frequencies obtained via more expansive random sampling. One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012]. Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem (e.g. [@Berg2012]).

@@

The negative consequences of excessive length sampling on abundance at age estimates are most apparent under the low set density scenarios. The precision of the estimates are also the poorest when fewer sets are conducted. Sets, however, are the most costly sample to obtain and, as such, the set densities of fisheries-independent surveys tend to be on the lower end of the scenarios presented in this paper. The set density of the multi-species survey conducted by Fisheries and Oceans Canada in the Newfoundland region, for instance, ranges between 0.001 and 0.002 sets / km^2^. Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively easy to obtain. Again, using Fisheries and Oceans Canada sampling protocol from the Newfoundland region as an example, it is not uncommon for length sampling effort to be capped at 500 measurements per set. The results presented in this paper paradoxically suggest that better abundance at age estimates are obtained by lowering the length sampling cap. Results also indicate that reductions to length sampling effort would not significantly impact abundance at length estimates. Nonetheless, it is hard to be prescriptive because this simulation is far from a perfect reflection of reality and it focuses on one case study. Further research is required on the consequences of some of the assumptions of this simulation as well as the interaction between age-specific clustering, length-stratified sampling and the age-length key. Ideally, the simulation would also include multiple species with different life-histories and distributions, and also integrate a cost component to assess the trade-offs between information and cost. Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key (e.g. [@Berg2012]).

@@

Standard notation for use in the analysis of stratified-random survey data (modified from [@Smith1981]):
